### 2022-11-20 analysis.Rmd
analysis

---
title: "PHQ-4-R: Adaptation of the 4-item screening for Anxiety and Depression for a subclinical sensitivity"
subtitle: "Analysis"
author: "Dominique Makowski et al."
output:
  html_document:
    toc: true
    toc_depth: 2
    theme: paper
    code_folding: hide
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---

```{r, echo = FALSE, warning=FALSE, message=FALSE}
options(
  digits = 3,
  mc.cores = 4,
  brms.algorithm = "sampling",
  brms.backend = "cmdstanr",
  dplyr.summarise.inform = FALSE,
  knitr.kable.NA = ""
)

fig.width <- see::golden_ratio(7)
fig.height <- 7

knitr::opts_chunk$set(
  collapse = TRUE,
  dpi = 450,
  fig.path = "./figures/",
  fig.width = fig.width,
  fig.height = fig.height
)
```

See [**tutorial**](https://bookdown.org/bean_jerry/using_r_for_social_work_research/item-response-theory.html) on Item Response Theory (IRT).

# Data

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggside)
library(easystats)
library(patchwork)
library(mirt)

df <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGameReliability/main/data/preprocessed_questionnaires.csv") |> 
  select(starts_with("Item_PHQ4")) |> 
  datawizard::remove_empty_rows()

names(df) <- str_remove_all(names(df), "Item_PHQ4_")
```

# Depression

```{r message=FALSE, warning=FALSE}
# s <- 'Anxiety = 1,2
#       Depression = 3,4
#       COV = Anxiety*Depression'
# mod1 <- mirt(df, model = mirt.model(s), itemtype = 'graded', SE = TRUE)

data <- select(df, contains("Depression"))

mod1 <- mirt(data, model = 1, itemtype = 'graded', SE = TRUE)

# M2(mod1, type = "C2", calcNULL = FALSE)

itemfit(mod1)

# IRT parameters
IRT_parms <- coef(mod1, IRTpars = TRUE, simplify = TRUE)
IRT_parms$items

# Factor
summary(mod1)

# Plots
plot(mod1, type = 'trace', theta_lim = c(-3, 3))
plot(mod1, type = 'infotrace', theta_lim = c(-3, 3))
plot(mod1, type = 'score', theta_lim = c(-3, 3))
plot(mod1, type = 'infoSE', theta_lim = c(-3, 3))
plot(mod1, type = 'rxx', theta_lim = c(-3, 3))
```

## Category Characteristic Curves (CRC) and Item Information Curves

Typically, an optimally informative item will have a large location and broad category coverage (as indicated by location parameters) over theta.

```{r message=FALSE, warning=FALSE}
crc <- function(mod, data, x = c(-1.25, 2.25)) {
  Theta <- matrix(seq(-3, 3, by = .01))
  rez <- data.frame()
  for(i in 1:2) {
    dat <- as.data.frame(probtrace(extract.item(mod1, i), Theta))
    dat$Theta <- Theta[,1]
    dat$Information <- normalize(iteminfo(extract.item(mod1, i), Theta))
    dat <- pivot_longer(dat, -one_of(c("Theta", "Information")), names_to = "Answer", values_to = "Probability")
    dat$Item <- names(data)[i]
    rez <- rbind(rez, dat)
  }

  rez <- rez |> 
    mutate(
      Answer = case_when(
      Answer == "P.1" ~ "Not at all",
      Answer == "P.2" ~ "Once or twice",
      Answer == "P.3" ~ "Several days",
      Answer == "P.4" ~ "More than half the days",
      TRUE ~ "Nearly every day"
      )) |> 
    mutate(
      Answer = fct_relevel(Answer, c("Not at all", "Once or twice", "Several days", "More than half the days",  "Nearly every day")))

  p <- rez |> 
    ggplot(aes(x = Theta, y = Probability)) +
    geom_line(aes(y = Information), linetype = "dotted", color = "grey") + 
    geom_line(aes(color = Answer)) + 
    geom_vline(xintercept = x, linetype = "dashed") +
    facet_grid(~Item) + 
    theme_modern() 
  list(p = p, rez = rez)
}

out <- crc(mod1, data, x = c(-1.25, 2.25))
out$p + labs(x = expression(Depression~(theta))) 
```

## Normalized scoring

```{r message=FALSE, warning=FALSE}
item4 <- out$rez |> 
  filter(Item == "Depression_4") |> 
  filter(Theta >= -1.5 & Theta <= 2.25) |> 
  mutate(Theta = normalize(Theta))

scores <- item4 |> 
  group_by(Answer) |> 
  filter(Probability == max(Probability)) |> 
  ungroup()

item4 |> 
  ggplot(aes(x = Theta, y = Probability, color = Answer)) +
  geom_line() + 
  theme_modern() 
```

# Anxiety

```{r message=FALSE, warning=FALSE}
data <- select(df, contains("Anxiety"))

m_anxiety <- mirt(data, model = 1, itemtype = 'graded', SE = TRUE)

# M2(m_anxiety, type = "C2", calcNULL = FALSE)

itemfit(m_anxiety)

# IRT parameters
IRT_parms <- coef(m_anxiety, IRTpars = TRUE, simplify = TRUE)
IRT_parms$items

# Factor
summary(m_anxiety)

# Plots
plot(m_anxiety, type = 'score', theta_lim = c(-3, 3))
plot(m_anxiety, type = 'infoSE', theta_lim = c(-3, 3))
plot(m_anxiety, type = 'rxx', theta_lim = c(-3, 3))
```

## Category Characteristic Curves (CRC) and Item Information Curves

```{r message=FALSE, warning=FALSE}
out <- crc(m_anxiety, data, x = c(-1.25, 2.25))
out$p + labs(x = expression(Anxiety~(theta))) 
```