### 2023-01-17 analysis.Rmd
update JAD

---
title: "PHQ-4-R: Adaptation of the 4-item screening for Anxiety and Depression for a subclinical sensitivity"
subtitle: "Analysis"
author: "Dominique Makowski et al."
output:
  html_document:
    toc: true
    toc_depth: 2
    theme: paper
    code_folding: hide
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---

```{r, echo = FALSE, warning=FALSE, message=FALSE}
options(
  digits = 3,
  mc.cores = 4,
  brms.algorithm = "sampling",
  brms.backend = "cmdstanr",
  dplyr.summarise.inform = FALSE,
  knitr.kable.NA = ""
)

fig.width <- see::golden_ratio(7)
fig.height <- 7
set.seed(333)

knitr::opts_chunk$set(
  collapse = TRUE,
  dpi = 450,
  fig.path = "./figures/",
  fig.width = fig.width,
  fig.height = fig.height
)
```

<!-- See [**tutorial**](https://bookdown.org/bean_jerry/using_r_for_social_work_research/item-response-theory.html) on Item Response Theory (IRT). -->

<!-- https://aidenloe.github.io/introToIRT.html -->

# Data

The data has been collected during the study of . The dataset is available in open-access at ...

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggside)
library(easystats)
library(patchwork)
library(mirt)

df <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGameReliability/main/data/preprocessed_questionnaires.csv") |>
# df <- read.csv("../IllusionGameReliability/data/preprocessed_questionnaires.csv") |> 
  select(Sex, Age, starts_with("Item_PHQ4")) |> 
  filter(!is.na(Item_PHQ4_Anxiety_1))

names(df) <- str_remove_all(names(df), "Item_PHQ4_")
```

# Participants

```{r message=FALSE, warning=FALSE}
report::report_participants(df)
```

# Distribution

```{r message=FALSE, warning=FALSE}
add_labels <- function(x) {
  x <- case_when(
      x == 0 ~ "Not at all",
      x == 1 ~ "Once or twice",
      x == 2 ~ "Several days",
      x == 3 ~ "More than half the days",
      TRUE ~ "Nearly every day"
    )
  fct_relevel(x, c("Not at all", "Once or twice", "Several days", "More than half the days",  "Nearly every day"))
}
```

```{r distribution, message=FALSE, warning=FALSE, fig.width=fig.height}
df <- select(df, -Sex, -Age)

data <- df |> 
  pivot_longer(everything(), names_to = "Item", values_to = "Answer") |> 
  group_by(Item, Answer) |> 
  summarise(n = n() / nrow(df)) |> 
  mutate(
    Facet = str_split_fixed(Item, "_", 2)[, 1],
    # Item = str_split_fixed(Item, "_", 2)[,2],
    Answer = add_labels(Answer),
    Item = case_when(
      Item == "Anxiety_1" ~ "A1: Feeling nervous, anxious or on edge",
      Item == "Anxiety_2" ~ "A2: Not being able to stop or control worrying",
      Item == "Depression_3" ~ "D1: Little interest or pleasure in doing things",
      TRUE ~ "D2: Feeling down, depressed, or hopeless"
    ))

p1 <- data |> 
  ggplot(aes(x = Answer, fill=Facet)) +
  geom_bar(aes(y = n), stat = "identity") +
  scale_y_continuous(labels=scales::percent, expand = c(0, 0)) +
  scale_fill_manual(values = c("Depression" = "#3F51B5", "Anxiety" = "#E91E63"), guide = "none") +
  labs(y = "Proportion of Answers") +
  theme_modern(axis.title.space = 10) +
  theme(axis.text.x = element_text(angle = 45, hjust=1),
        axis.title.x = element_blank(),
        plot.title = element_text(face = "bold"),
        strip.text = element_text(size=rel(0.7)),
        strip.background=element_rect(fill="#EEEEEE", colour="white")) +
  facet_wrap(~Item)
p1 + ggtitle("Proportion of answers per item")
```

The "Once or twice" answer was selected in `r insight::format_percent(mean(filter(data, Answer == "Once or twice")$n))`

```{r joint_prevalence, message=FALSE, warning=FALSE, fig.width=fig.width}
p1b <- rbind(
  data.frame(Scores = paste0(df$Anxiety_1, "_", df$Anxiety_2),
             Facet = "Anxiety"),
  data.frame(Scores = paste0(df$Depression_3, "_", df$Depression_4),
             Facet = "Depression")
) |> 
  group_by(Facet, Scores) |> 
  summarize(n = n() / nrow(df)) |> 
  separate(Scores, into = c("Q1", "Q2")) |> 
  mutate(Label = ifelse(Q1 == Q2, format_percent(n), "")) |> 
  mutate(Q1 = add_labels(as.numeric(Q1)),
         Q2 = add_labels(as.numeric(Q2))) |> 
  ggplot(aes(x = Q1, y = Q2)) +
  geom_tile(aes(fill = n)) +
  geom_text(aes(label = Label), size=2, color="white") +
  scale_y_discrete(expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  scale_fill_gradientn(colors=c("white", "#2196F3", "#3F51B5", "#673AB7"), labels = scales::percent) +
  labs(fill = "% Participants", x = "Item 1", y = "Item 2") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust=1),
        panel.grid.major = element_blank(),
        plot.title = element_text(face = "bold", hjust=0),
        strip.background=element_rect(fill="#EEEEEE", colour="white"),
        strip.text = element_text(face = "bold")) +
  facet_wrap(~Facet) 
p1b + ggtitle("Joint prevalence of responses")
```

# Anxiety

## Descriptive
```{r message=FALSE, warning=FALSE}
df_anx <- select(df, contains("Anxiety"))

performance::cronbachs_alpha(df_anx)
```

## Model

```{r message=FALSE, warning=FALSE}
m_anxiety <- mirt(df_anx, model = 1, itemtype = 'graded', SE = TRUE)

m_anxiety

# Factor loadings
summary(m_anxiety)

# Alpha
a <- coef(m_anxiety)
a <- data.frame(
  Item = c("Anxiety 1", "Anxiety 2"),
  Discrimination = c(paste0(format_value(a$Anxiety_1[1, 1]), ", ", format_ci(a$Anxiety_1[2, 1], a$Anxiety_1[3, 1])), paste0(format_value(a$Anxiety_2[1, 1]), ", ", format_ci(a$Anxiety_2[2, 1], a$Anxiety_2[3, 1]))
))
knitr::kable(a)
```

```{r message=FALSE, warning=FALSE, eval=FALSE}
# Plots
plot(m_anxiety, type = 'trace', theta_lim = c(-3, 3))
plot(m_anxiety, type = 'infotrace', theta_lim = c(-3, 3))
plot(m_anxiety, type = 'score', theta_lim = c(-3, 3))
plot(m_anxiety, type = 'infoSE', theta_lim = c(-3, 3))
plot(m_anxiety, type = 'rxx', theta_lim = c(-3, 3))
```

## Category Characteristic Curves (CRC) and Item Information Curves

Typically, an optimally informative item will have a large location and broad category coverage (as indicated by location parameters) over theta.

```{r CRC_anxiety, message=FALSE, warning=FALSE}
crc <- function(mod, data) {
  Theta <- matrix(seq(-3.5, 3.5, by = .01))
  rez <- data.frame()
  for(i in 1:2) {
    dat <- as.data.frame(probtrace(extract.item(mod, i), Theta))
    dat$Theta <- Theta[,1]
    dat$Information <- normalize(iteminfo(extract.item(mod, i), Theta))
    dat <- pivot_longer(dat, -one_of(c("Theta", "Information")), names_to = "Answer", values_to = "Probability")
    dat$Item <- names(data)[i]
    rez <- rbind(rez, dat)
  }

  rez <- rez |>
    mutate(
      Answer = case_when(
      Answer == "P.1" ~ "Not at all",
      Answer == "P.2" ~ "Once or twice",
      Answer == "P.3" ~ "Several days",
      Answer == "P.4" ~ "More than half the days",
      TRUE ~ "Nearly every day"
      ),
      Answer = fct_relevel(Answer, c("Not at all", "Once or twice", "Several days", "More than half the days",  "Nearly every day")),
    Item = case_when(
      Item == "Anxiety_1" ~ "A1: Feeling nervous, anxious or on edge",
      Item == "Anxiety_2" ~ "A2: Not being able to stop or control worrying",
      Item == "Depression_3" ~ "D1: Little interest or pleasure in doing things",
      TRUE ~ "D2: Feeling down, depressed, or hopeless"
    ))

  # Get close to 0.95
  i1 <- rez[rez$Item == unique(rez$Item)[1],]
  i2 <- rez[rez$Item == unique(rez$Item)[2],]
  minmax <- rbind(
    rbind(
      i1[i1$Answer == "Not at all", ][which.min(abs(i1[i1$Answer == "Not at all", ]$Probability - 0.95)), ],
      i1[i1$Answer == "Nearly every day", ][which.min(abs(i1[i1$Answer == "Nearly every day", ]$Probability - 0.95)), ]
      ),
    rbind(
      i2[i2$Answer == "Not at all", ][which.min(abs(i2[i2$Answer == "Not at all", ]$Probability - 0.95)), ],
      i2[i2$Answer == "Nearly every day", ][which.min(abs(i2[i2$Answer == "Nearly every day", ]$Probability - 0.95)), ]
      )
  )

  p <- rez |>
    ggplot(aes(x = Theta, y = Probability)) +
    geom_line(aes(y = Information), linetype = "dotted", color = "grey") +
    geom_line(aes(color = Answer), size=1) +
    # geom_vline(data = minmax, aes(xintercept = Theta), linetype = "dashed") +
    scale_y_continuous(labels=scales::percent, expand = c(0, 0.005)) +
    scale_color_flat_d("rainbow") +
    facet_grid(~Item) +
    theme_modern(axis.title.space = 5) +
    theme(strip.background=element_rect(fill="#EEEEEE", colour="white"))
  list(p = p, rez = rez, minmax=minmax)
}

out <- crc(m_anxiety, df_anx)
p2a <- out$p + labs(x = expression(Anxiety~(theta)))
p2a
```

## Normalized scoring

```{r message=FALSE, warning=FALSE}
normalize_scores <- function(data, minmax) {

  minmax <- minmax[minmax$Theta %in% range(minmax$Theta), ]

  item <- data.frame()
  scores <- data.frame()
  for(i in unique(data$Item)) {
    item <- data[data$Item == i, ] |>
      filter(Theta >= min(minmax$Theta) & Theta <= max(minmax$Theta)) |>
      mutate(Theta_n = normalize(Theta)) |> 
      rbind(item)

    scores <- item[item$Item == i, ] |>
      group_by(Answer) |>
      filter(Probability == max(Probability)) |>
      ungroup() |>
      mutate(label = insight::format_value(Theta_n)) |> 
      rbind(scores)
  }

  item |>
    ggplot(aes(x = Theta, y = Probability, color = Answer)) +
    geom_line(size = 1) +
    geom_segment(data=scores, aes(x=Theta, xend = Theta, y=0, yend=Probability, color = Answer), linetype = "dashed") +
    geom_label(data=scores, aes(x=Theta, y=Probability, label=label), show.legend = FALSE) +
    scale_y_continuous(labels=scales::percent, expand = c(0, 0.01)) +
    scale_color_viridis_d(option = "D") +
    facet_grid(~Item) +
    theme_modern(axis.title.space = 5) +
    theme(strip.background=element_rect(fill="#EEEEEE", colour="white"))
}

p3a <- normalize_scores(data=out$rez, out$minmax) + labs(x = expression(Anxiety~(theta)))
p3a 
```

# Depression

## Descriptive
```{r message=FALSE, warning=FALSE}
df_dep <- select(df, contains("Depression"))

performance::cronbachs_alpha(df_dep)
```

## Model

```{r message=FALSE, warning=FALSE}
m_depression <- mirt(df_dep, model = 1, itemtype = 'graded', SE = TRUE)

m_depression

# Factor loadings
summary(m_depression)

# Alpha
a2 <- coef(m_depression)
a2 <- data.frame(
  Item = c("Depression 1", "Depression 2"),
  Discrimination = c(paste0(format_value(a2$Depression_3[1, 1]), ", ", format_ci(a2$Depression_3[2, 1], a2$Depression_3[3, 1])), paste0(format_value(a2$Depression_4[1, 1]), ", ", format_ci(a2$Depression_4[2, 1], a2$Depression_4[3, 1]))
))
knitr::kable(a2)
```

## Category Characteristic Curves (CRC) and Item Information Curves

```{r message=FALSE, warning=FALSE}
out <- crc(m_depression, df_dep)
p2b <- out$p + labs(x = expression(Depression~(theta)))
p2b
```

## Normalized scoring

```{r message=FALSE, warning=FALSE}
p3b <- normalize_scores(out$rez, out$minmax) + labs(x = expression(Depression~(theta)))
p3b 
```

# Figures

```{r message=FALSE, warning=FALSE}
fig1 <- wrap_elements(wrap_elements(p1 + plot_annotation(title = "A. Proportion of answers per item", theme = list(plot.title = element_text(face = "bold")))) / 
                        wrap_elements(p1b + plot_annotation(title = "B. Joint prevalence of responses", theme = list(plot.title = element_text(face = "bold"))))) + 
  wrap_elements(p2a / p2b + plot_annotation(title = "C. Information curves per response type", theme = list(plot.title = element_text(face = "bold"))) + plot_layout(guides = "collect")) + plot_layout(widths = c(1, 1.5))

ggsave("figures/figure1.png", fig1, width = fig.width*1.43, height=fig.height*1.43)

fig2 <- p3a / p3b + plot_annotation(title = "Normalized Scoring", theme = list(plot.title = element_text(face = "bold"))) + plot_layout(guides = "collect")

ggsave("figures/figure2.png", fig2, width = fig.width*1.43, height=fig.height*1.43)
```

# Scores

```{r message=FALSE, warning=FALSE}
data <- sapply(cbind(df_anx, df_dep), function(x) {
  case_when(
      x == 0 ~ "Not at all",
      x == 1 ~ "Once or twice",
      x == 2 ~ "Several days",
      x == 3 ~ "More than half the days",
      TRUE ~ "Nearly every day"
    )
}) |> 
  as.data.frame()

source("score_PHQ4.R")

data <- score_PHQ4(A1=data$Anxiety_1, A2=data$Anxiety_2, D1=data$Depression_3, D2=data$Depression_4, method="basic") |> 
  datawizard::data_addsuffix("_Basic") |> 
  cbind(
    score_PHQ4(A1=data$Anxiety_1, A2=data$Anxiety_2, D1=data$Depression_3, D2=data$Depression_4, method="normalized") |> 
      datawizard::data_addsuffix("_Normalized")) |> 
  select(matches("Depression|Anxiety")) |> 
  pivot_longer(everything()) |> 
  separate(name, into = c("Dimension", "Scoring"))

data |> 
  ggplot(aes(x=value)) + 
  geom_area(data=estimate_density(data, at = c("Dimension", "Scoring"), method = "KernSmooth") |> 
              group_by(Scoring) |> 
              normalize(select = "y"),
            aes(x = x, y = y, fill=Dimension), alpha = 0.05, position="identity") +
  geom_hline(yintercept = 0.5, linetype = "dotted") +
  stat_ecdf(aes(color=Dimension), geom = "step", size=1) +
  facet_wrap(~Scoring, scales = "free") +
  scale_y_continuous(labels=scales::percent, expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_fill_manual(values = c("Depression" = "#3F51B5", "Anxiety" = "#E91E63"), guide = "none") +
  scale_color_manual(values = c("Depression" = "#3F51B5", "Anxiety" = "#E91E63")) +
  theme_modern(axis.title.space = 5) +
  labs(title = "Distribution and Cumulative Density of the Sample", x = "Score", y = "Cumulative Proportion")
```

# Correlation

```{r message=FALSE, warning=FALSE}
df$Depression <- mirt::fscores(m_depression)[, 1]
df$Anxiety <- mirt::fscores(m_anxiety)[, 1]

correlation::correlation(df) |>
  summary() |> 
  export_table(caption = "Correlation Matrix")
```

<!-- # Multifactor model -->

<!-- ```{r message=FALSE, warning=FALSE, eval=FALSE} -->
<!-- s <- 'Anxiety = 1,2 -->
<!--       Depression = 3,4 -->
<!--       COV = Anxiety*Depression' -->
<!-- mod <- mirt(df, model = mirt.model(s), itemtype = 'graded', SE = TRUE) -->
<!-- mod <- mirt(df, model = 2, itemtype = 'graded', SE = TRUE) -->

<!-- summary(mod) -->

<!-- # Plots -->
<!-- plot(mod, type = 'trace', theta_lim = c(-3, 3)) -->
<!-- plot(mod, type = 'infotrace', theta_lim = c(-3, 3)) -->
<!-- plot(mod, type = 'score', theta_lim = c(-3, 3)) -->

<!-- Theta <- as.matrix(expand.grid(seq(-3, 3, by = 1), seq(-3, 3, by = 1))) -->
<!-- rez <- data.frame() -->
<!-- for(i in 1:4) { -->
<!--   dat <- as.data.frame(probtrace(extract.item(mod, i), Theta)) -->
<!--   dat$Theta_F1 <- Theta[, 1] -->
<!--   dat$Theta_F2 <- Theta[, 2] -->
<!--   # dat$Information <- normalize(iteminfo(extract.item(mod, i), Theta)) -->
<!--   dat <- pivot_longer(dat, -one_of(c("Theta_F1", "Theta_F2")), names_to = "Answer", values_to = "Probability") -->
<!--   dat$Item <- names(df)[i] -->
<!--   rez <- rbind(rez, dat) -->
<!-- } -->

<!-- rez <- rez |> -->
<!--   separate(Item, into = c("Dimension", "Item_Number")) |> -->
<!--   mutate( -->
<!--     Answer = case_when( -->
<!--     Answer == "P.1" ~ "Not at all", -->
<!--     Answer == "P.2" ~ "Once or twice", -->
<!--     Answer == "P.3" ~ "Several days", -->
<!--     Answer == "P.4" ~ "More than half the days", -->
<!--     TRUE ~ "Nearly every day" -->
<!--     )) |> -->
<!--   mutate( -->
<!--     Answer = fct_relevel(Answer, c("Not at all", "Once or twice", "Several days", "More than half the days",  "Nearly every day"))) -->

<!-- rez |> -->
<!--   ggplot(aes(x = Theta_F1, y = Probability)) + -->
<!--   # geom_line(aes(y = Information), linetype = "dotted", color = "grey") + -->
<!--   geom_line(aes(color = Answer)) + -->
<!--   geom_vline(xintercept = c(-1.25, 2.25), linetype = "dashed") + -->
<!--   facet_grid(Dimension~Item_Number) + -->
<!--   theme_modern() -->
<!-- ``` -->