### 2024-02-12 study2/analysis/2_analysis.qmd

---
title: "PHQ4R - Data Analysis (Study 2)"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: true
---

## Data Preparation

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
```

```{r}
df <- read.csv("../data/data.csv") |> 
  mutate(
    PHQ4_Condition = fct_relevel(PHQ4_Condition, "PHQ4 - Original"),
    PHQ4_Condition = fct_recode(PHQ4_Condition, "Original" = "PHQ4 - Original", "Revised" = "PHQ4 - Revised"),
    BDI2_Category = ifelse(BDI2_Total <= 19, "Mild", "Moderate+"),
    STAI5_Category = ifelse(STAI5_General < 2, "Mild", "Moderate+"),
  )
```

## Depression

### Whole Sample

```{r}
model <- lm(BDI2_Total ~ PHQ4_Depression * PHQ4_Condition, data = df)

parameters::parameters(model) |> 
  display()
```

```{r}
df |> 
  ggplot(aes(x = PHQ4_Depression, y = BDI2_Total), width=0.1) +
  geom_jitter(aes(color = PHQ4_Condition), width=0.1) +
  geom_smooth(aes(color = PHQ4_Condition), method = "lm") +
  theme_bw() +
  ggside::geom_xsidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
  ggside::geom_ysidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_ysidex_continuous(expand = c(0, 0)) +
  ggside::theme_ggside_void() 
```

### Subsamples

```{r}
model <- lm(BDI2_Total ~ BDI2_Category / (PHQ4_Depression * PHQ4_Condition), data = df)

parameters::parameters(model) |> 
  display()

# model <- rstanarm::stan_glm(BDI2_Total ~ BDI2_Category / (PHQ4_Depression * PHQ4_Condition), data = df)
# parameters::parameters(model)
# bayestestR::bayesfactor_parameters(model)
```

```{r}
df |> 
  ggplot(aes(x = PHQ4_Depression, y = BDI2_Total)) +
  geom_jitter(aes(color = PHQ4_Condition), width=0.1) +
  geom_smooth(aes(color = PHQ4_Condition, linetype=BDI2_Category), method = "lm") +
  theme_bw() +
  ggside::geom_xsidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
  ggside::geom_ysidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_ysidex_continuous(expand = c(0, 0)) +
  ggside::theme_ggside_void() 
```

### General Score

```{r}
lm(BDI2_Total ~ PHQ4_Total * PHQ4_Condition, data = df) |> 
  parameters::parameters() |> 
  display()

lm(BDI2_Total ~ BDI2_Category / (PHQ4_Total * PHQ4_Condition), data = df) |> 
  parameters::parameters() |> 
  display()
```

```{r}
m0 <- lm(BDI2_Total ~ PHQ4_Depression, data = filter(df, PHQ4_Condition == "Original"))
m1 <- lm(BDI2_Total ~ PHQ4_Depression, data = filter(df, PHQ4_Condition == "Revised"))

performance::compare_performance(m0, m1)
```

## Anxiety

### Whole Sample

```{r}
model <- lm(STAI5_General ~ PHQ4_Anxiety * PHQ4_Condition, data = df)

parameters::parameters(model) |> 
  display()
```

```{r}
df |> 
  ggplot(aes(x = PHQ4_Anxiety, y = STAI5_General)) +
  geom_jitter(aes(color = PHQ4_Condition), width=0.1) +
  geom_smooth(aes(color = PHQ4_Condition), method = "lm") +
  theme_bw() +
  ggside::geom_xsidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
  ggside::geom_ysidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_ysidex_continuous(expand = c(0, 0)) +
  ggside::theme_ggside_void() 
```

### Subsamples

```{r}
model <- lm(STAI5_General ~ STAI5_Category / (PHQ4_Anxiety * PHQ4_Condition), data = df)

parameters::parameters(model) |> 
  display()
```

```{r}
df |> 
  ggplot(aes(x = PHQ4_Anxiety, y = STAI5_General)) +
  geom_jitter(aes(color = PHQ4_Condition), width=0.1) +
  geom_smooth(aes(color = PHQ4_Condition, linetype=STAI5_Category), method = "lm") +
  theme_bw() +
  ggside::geom_xsidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
  ggside::geom_ysidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_ysidex_continuous(expand = c(0, 0)) +
  ggside::theme_ggside_void() 
```

### General Score

```{r}
lm(STAI5_General ~ PHQ4_Total * PHQ4_Condition, data = df) |> 
  parameters::parameters() |> 
  display()

lm(STAI5_General ~ BDI2_Category / (PHQ4_Total * PHQ4_Condition), data = df) |> 
  parameters::parameters() |> 
  display()
```

## Correlation Difference

### Test

```{r}
dat <- df |> 
  group_by(PHQ4_Condition) |>
  summarize(
    N = n(),
    r_depression = cor.test(BDI2_Total, PHQ4_Depression)$estimate,
    r_anxiety = cor.test(STAI5_General, PHQ4_Anxiety)$estimate,
  )

psych::r.test(n=dat$N[1], n2=dat$N[2], r12=dat$r_depression[1], r34=dat$r_depression[2])
psych::r.test(n=dat$N[1], n2=dat$N[2], r12=dat$r_anxiety[1], r34=dat$r_anxiety[2])
```

### Bootstrapping

#### Whole Sample

```{r}
bootstrap_diffs <- function(df_original, df_revised) {
  dat <- data.frame()
  for(i in 1:1000) {
    newdf_ori <- df_original[sample(nrow(df_original), replace = TRUE), ]
    newdf_rev <- df_revised[sample(nrow(df_revised), replace = TRUE), ]

    r_depression_ori = cor.test(newdf_ori$BDI2_Total, newdf_ori$PHQ4_Depression)$estimate
    r_depression_rev = cor.test(newdf_rev$BDI2_Total, newdf_rev$PHQ4_Depression)$estimate
    r_anxiety_ori = cor.test(newdf_ori$STAI5_General, newdf_ori$PHQ4_Anxiety)$estimate
    r_anxiety_rev = cor.test(newdf_rev$STAI5_General, newdf_rev$PHQ4_Anxiety)$estimate

    dat <- rbind(dat,
      data.frame(
        d_r_depression = r_depression_rev - r_depression_ori,
        d_r_anxiety = r_anxiety_rev - r_anxiety_ori,
        iteration = i
      )
    )
  }
  dat
}

dat <- bootstrap_diffs(
  filter(df, PHQ4_Condition == "Original"),
  filter(df, PHQ4_Condition == "Revised"))

dat |> 
  pivot_longer(cols = c(d_r_depression, d_r_anxiety)) |>
  ggplot(aes(x = value)) +
  geom_density(aes(fill = name), alpha = 0.5) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_bw() +
  labs(x = "Difference in Correlation", y = "Density") 
```

#### Subsamples

```{r}
dat <- data.frame()
for(cat in c("Mild", "Moderate+")) {
  dat <- rbind(
    dat,
    bootstrap_diffs(
      filter(df, PHQ4_Condition == "Original", BDI2_Category == cat),
      filter(df, PHQ4_Condition == "Revised", BDI2_Category == cat)) |> 
      mutate(Category = paste("Depression -", cat)),
    bootstrap_diffs(
      filter(df, PHQ4_Condition == "Original", STAI5_Category == cat),
      filter(df, PHQ4_Condition == "Revised", STAI5_Category == cat)) |> 
      mutate(Category = paste("Anxiety -", cat)) 
  )
}

dat |> 
  separate(Category, into = c("Type", "Category"), sep = " - ") |>
  pivot_longer(cols = c(d_r_depression, d_r_anxiety)) |>
  ggplot(aes(x = value)) +
  geom_density(aes(fill = name), alpha = 0.5) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_bw() +
  labs(x = "Difference in Correlation", y = "Density") +
  facet_grid(Category~Type)
```

## Predictive Value

### Mood Disorder

```{r}
glm(MoodDisorder ~ PHQ4_Total * PHQ4_Condition, data = df, family="binomial") |> 
  parameters::parameters() |> 
  display()

glm(MoodDisorder ~ PHQ4_Depression * PHQ4_Condition, data = df, family="binomial") |> 
  parameters::parameters() |> 
  display()

glm(MoodDisorder ~ PHQ4_Anxiety * PHQ4_Condition, data = df, family="binomial") |> 
  parameters::parameters() |> 
  display()
```

### MDD

```{r}
glm(Disorder_MDD ~ PHQ4_Total * PHQ4_Condition, data = df, family="binomial") |> 
  parameters::parameters() |> 
  display()

glm(Disorder_MDD ~ PHQ4_Depression * PHQ4_Condition, data = df, family="binomial") |> 
  parameters::parameters() |> 
  display()

glm(Disorder_MDD ~ PHQ4_Anxiety * PHQ4_Condition, data = df, family="binomial") |> 
  parameters::parameters() |> 
  display()
```

### GAD

```{r}
glm(Disorder_GAD ~ PHQ4_Total * PHQ4_Condition, data = df, family="binomial") |> 
  parameters::parameters() |> 
  display()

glm(Disorder_GAD ~ PHQ4_Depression * PHQ4_Condition, data = df, family="binomial") |> 
  parameters::parameters() |> 
  display()

glm(Disorder_GAD ~ PHQ4_Anxiety * PHQ4_Condition, data = df, family="binomial") |> 
  parameters::parameters() |> 
  display()
```

## Modified - Depression

### Whole Sample

```{r}
model <- lm(BDI2_Total ~ Expe_PHQ4_Depression * PHQ4_Condition, data = df)

parameters::parameters(model) |> 
  display()
```

```{r}
df |> 
  ggplot(aes(x = Expe_PHQ4_Depression, y = BDI2_Total), width=0.1) +
  geom_jitter(aes(color = PHQ4_Condition), width=0.1) +
  geom_smooth(aes(color = PHQ4_Condition), method = "lm") +
  theme_bw() +
  ggside::geom_xsidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
  ggside::geom_ysidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_ysidex_continuous(expand = c(0, 0)) +
  ggside::theme_ggside_void() 
```

### Subsamples

```{r}
model <- lm(BDI2_Total ~ BDI2_Category / (Expe_PHQ4_Depression * PHQ4_Condition), data = df)

parameters::parameters(model) |> 
  display()
```

```{r}
df |> 
  ggplot(aes(x = Expe_PHQ4_Depression, y = BDI2_Total)) +
  geom_jitter(aes(color = PHQ4_Condition), width=0.1) +
  geom_smooth(aes(color = PHQ4_Condition, linetype=BDI2_Category), method = "lm") +
  theme_bw() +
  ggside::geom_xsidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
  ggside::geom_ysidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_ysidex_continuous(expand = c(0, 0)) +
  ggside::theme_ggside_void() 
```

## Modified - Anxiety

### Whole Sample

```{r}
model <- lm(STAI5_General ~ Expe_PHQ4_Anxiety * PHQ4_Condition, data = df)

parameters::parameters(model) |> 
  display()
```

```{r}
df |> 
  ggplot(aes(x = Expe_PHQ4_Anxiety, y = STAI5_General)) +
  geom_jitter(aes(color = PHQ4_Condition), width=0.1) +
  geom_smooth(aes(color = PHQ4_Condition), method = "lm") +
  theme_bw() +
  ggside::geom_xsidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
  ggside::geom_ysidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_ysidex_continuous(expand = c(0, 0)) +
  ggside::theme_ggside_void() 
```

### Subsamples

```{r}
model <- lm(STAI5_General ~ STAI5_Category / (Expe_PHQ4_Anxiety * PHQ4_Condition), data = df)

parameters::parameters(model) |> 
  display()
```

```{r}
df |> 
  ggplot(aes(x = Expe_PHQ4_Anxiety, y = STAI5_General)) +
  geom_jitter(aes(color = PHQ4_Condition), width=0.1) +
  geom_smooth(aes(color = PHQ4_Condition, linetype=STAI5_Category), method = "lm") +
  theme_bw() +
  ggside::geom_xsidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
  ggside::geom_ysidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_ysidex_continuous(expand = c(0, 0)) +
  ggside::theme_ggside_void() 
```