### 2024-02-19 study2/analysis/1_cleaning.qmd
update cleaning and data

---
title: "PHQ-4-R - Data Cleaning (Study 2)"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: true
---

## Data Preparation

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
```

```{r}
#| code-fold: false

df <- read.csv("../data/data_raw.csv")
```

The initial sample consisted of `r report::report_participants(df, age="Age", gender="Gender", education="Education")`.

### Recruitment History

```{r}
#| message: false

# Consecutive count of participants per day (as area)
df |>
  mutate(Date = as.Date(Date, format = "%d/%m/%Y")) |> 
  group_by(Date, Experimenter) |> 
  summarize(N = n()) |> 
  ungroup() |>
  complete(Date, Experimenter, fill = list(N = 0)) |> 
  group_by(Experimenter) |>
  mutate(N = cumsum(N)) |>
  ggplot(aes(x = Date, y = N)) +
  geom_area(aes(fill=Experimenter)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    title = "Recruitment History",
    x = "Date",
    y = "Total Number of Participants"
  ) +
  see::theme_modern()
```

### Compute Scores

#### PHQ-4

```{r}
#| echo: true
#| code-fold: show

# TODO: compute modified score
# Maybe use the score_PHQ4 function from the README to demo it here

recode_phq <- function(x) {
  case_when(
    x == 1 ~ 0.5,
    x == 2 ~ 1,
    x == 3 ~ 2,
    x == 4 ~ 3,
    .default = x
  )
}

df <- df |>
  mutate(
    PHQ4_Anxiety_1 = ifelse(PHQ4_Condition == "PHQ4 - Revised", recode_phq(PHQ4_Anxiety_1), PHQ4_Anxiety_1),
    PHQ4_Anxiety_2 = ifelse(PHQ4_Condition == "PHQ4 - Revised", recode_phq(PHQ4_Anxiety_2), PHQ4_Anxiety_2),
    PHQ4_Depression_3 = ifelse(PHQ4_Condition == "PHQ4 - Revised", recode_phq(PHQ4_Depression_3), PHQ4_Depression_3),
    PHQ4_Depression_4 = ifelse(PHQ4_Condition == "PHQ4 - Revised", recode_phq(PHQ4_Depression_4), PHQ4_Depression_4),
    PHQ4_Anxiety = PHQ4_Anxiety_1 + PHQ4_Anxiety_2,
    PHQ4_Depression = PHQ4_Depression_3 + PHQ4_Depression_4,
    PHQ4_Total = PHQ4_Anxiety + PHQ4_Depression
  )

df <- df |>
  mutate(
    Weighted_PHQ4_Anxiety_1 = ifelse(
      PHQ4_Condition == "PHQ4 - Revised",
      case_when(
        PHQ4_Anxiety_1 == 0 ~ 0,
        PHQ4_Anxiety_1 == 0.5 ~ 0.34,
        PHQ4_Anxiety_1 == 1 ~ 0.55,
        PHQ4_Anxiety_1 == 2 ~ 0.71,
        PHQ4_Anxiety_1 == 3 ~ 1,
        .default = PHQ4_Anxiety_1
      ), PHQ4_Anxiety_1
    ),
    Weighted_PHQ4_Anxiety_2 = ifelse(
      PHQ4_Condition == "PHQ4 - Revised",
      case_when(
        PHQ4_Anxiety_2 == 0 ~ 0,
        PHQ4_Anxiety_2 == 0.5 ~ 0.44,
        PHQ4_Anxiety_2 == 1 ~ 0.58,
        PHQ4_Anxiety_2 == 2 ~ 0.71,
        PHQ4_Anxiety_2 == 3 ~ 1,
        .default = PHQ4_Anxiety_2
      ), PHQ4_Anxiety_2
    ),
    Weighted_PHQ4_Depression_3 = ifelse(
      PHQ4_Condition == "PHQ4 - Revised",
      case_when(
        PHQ4_Depression_3 == 0 ~ 0,
        PHQ4_Depression_3 == 0.5 ~ 0.38,
        PHQ4_Depression_3 == 1 ~ 0.51,
        PHQ4_Depression_3 == 2 ~ 0.62,
        PHQ4_Depression_3 == 3 ~ 1,
        .default = PHQ4_Depression_3
      ), PHQ4_Depression_3
    ),
    Weighted_PHQ4_Depression_4 = ifelse(
      PHQ4_Condition == "PHQ4 - Revised",
      case_when(
        PHQ4_Depression_4 == 0 ~ 0,
        PHQ4_Depression_4 == 0.5 ~ 0.35,
        PHQ4_Depression_4 == 1 ~ 0.52,
        PHQ4_Depression_4 == 2 ~ 0.66,
        PHQ4_Depression_4 == 3 ~ 1,
        .default = PHQ4_Depression_4
      ), PHQ4_Depression_4
    ),
    Weighted_PHQ4_Anxiety = ifelse(
      PHQ4_Condition == "PHQ4 - Revised", datawizard::rescale(Weighted_PHQ4_Anxiety_1 + Weighted_PHQ4_Anxiety_2, range = c(0, 2), to = c(0, 6)), PHQ4_Anxiety_1 + PHQ4_Anxiety_2
    ),
    Weighted_PHQ4_Depression = ifelse(
      PHQ4_Condition == "PHQ4 - Revised", datawizard::rescale(Weighted_PHQ4_Depression_3 + Weighted_PHQ4_Depression_4, range = c(0, 2), to = c(0, 6)), PHQ4_Depression_3 + PHQ4_Depression_4
    ),
    Weighted_PHQ4_Total = Weighted_PHQ4_Anxiety + Weighted_PHQ4_Depression
  )

# df |>
#   select(starts_with("Weighted_PHQ4_"), PHQ4_Condition) |>
#   pivot_longer(cols = -PHQ4_Condition, names_to = "item", values_to = "score") |>
#   ggplot(aes(x = score)) +
#   geom_density(fill="#9C27B0", adjust=1/2)  +
#   facet_grid(item ~ PHQ4_Condition)
```

```{r}
df |>
  select(starts_with("PHQ4_")) |>
  pivot_longer(cols = matches("Depression_|Anxiety_"), names_to = "item", values_to = "score") |>
  ggplot(aes(x = as.factor(score))) +
  geom_bar(stat="count", fill="#9C27B0") +
  facet_grid(item ~ PHQ4_Condition) +
  labs(x = "Score", y = "N answers") +
  theme_bw()
```

#### STAI-5

```{r}
df |>
  select(matches("STAI5_"), -STAI5_Duration, -STAI5_Order) |>
  pivot_longer(cols = matches("STAI5_"), names_to = "item", values_to = "score") |>
  ggplot(aes(x = as.factor(score))) +
  geom_bar(stat="count", fill="#E91E63") +
  facet_wrap(~item) +
  labs(x = "Score", y = "N answers") +
  theme_bw()
```

```{r}
#| echo: true
#| code-fold: show

df <- df |> 
  mutate(STAI5_General = (STAI5_1 + STAI5_2 + STAI5_3 + STAI5_4 + STAI5_5) / 5)
```

<!-- df[is.na(df$STAI5_General), c("Participant")] -->

#### BDI-2

```{r}
df |>
  select(matches("BDI2_"), -BDI2_Duration, -BDI2_Order) |>
  pivot_longer(cols = matches("BDI2_"), names_to = "item", values_to = "score") |>
  ggplot(aes(x = as.factor(score))) +
  geom_bar(stat="count", fill="#2196F3") +
  facet_wrap(~item) +
  labs(x = "Score", y = "N answers") +
  theme_bw()
```

```{r}
#| echo: true
#| code-fold: show

df <- df |> 
  mutate(BDI2_Total = rowSums(select(df, matches("BDI2_"), -BDI2_Duration, -BDI2_Order)))
```

<!-- df[is.na(df$BDI2_Total), c("Participant")] -->

## Exclusions

### Experiment Duration

The experiment's median duration is `r insight::format_value(median(df$Experiment_Duration))` min (`r insight::format_ci(bayestestR::hdi(df$Experiment_Duration, ci=0.5))`).

```{r}
df |>
  mutate(Participant = fct_reorder(Participant, Experiment_Duration),
         Category = ifelse(Experiment_Duration > 30, "extra", "ok"),
         Duration = ifelse(Experiment_Duration > 30, 30, Experiment_Duration)) |>
  ggplot(aes(y = Participant, x = Duration)) +
  geom_vline(xintercept = median(df$Experiment_Duration), color = "red", linetype = "dashed") +
  geom_point(aes(color = Experimenter, shape = Category)) +
  geom_text(aes(color = Experimenter, label=ifelse(Experiment_Duration < 3, as.character(Participant), '')), hjust=-1, vjust=0) +
  scale_shape_manual(values = c("extra" = 3, ok = 19)) +
  guides(shape = "none") +
  ggside::geom_xsidedensity(aes(fill=Experimenter), color=NA, alpha=0.5) +
  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
  labs(
    title = "Experiment Completion Time",
    x = "Duration (in minutes)",
    y = "Participant"
  )  +
  theme_bw() +
  ggside::theme_ggside_void() +
  theme(ggside.panel.scale = .3,
        panel.border = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) 
```

### Response Coherence

```{r}
SD_per_dim <- function(x, dims="") {
  m <- matrix(nrow=nrow(x), ncol=0)
  for(s in dims) {
    m <- cbind(m, sapply(as.data.frame(t(x[grepl(s, names(x))])), sd))
  }
  m
}

df$BDI2_SD <- select(df, matches("BDI2_"), -BDI2_Duration, -BDI2_Order, -BDI2_Total) |> 
  SD_per_dim(dims=c("BDI2_")) |> 
  rowMeans()
df$STAI5_SD <- select(df, matches("STAI5_"), -STAI5_Duration, -STAI5_Order, -STAI5_General) |> 
  SD_per_dim(dims=c("STAI5_")) |> 
  rowMeans()
```

```{r}
outliers <- df |> 
  select(Participant, BDI2_SD, BDI2_Duration, STAI5_SD, STAI5_Duration) |> 
  standardize(robust=TRUE) |> 
  mutate(SD = rowSums(across(ends_with("SD"))), Duration = rowSums(across(ends_with("Duration"))))

# df |> 
#   ggplot(aes(x=BDI2_Duration, y=BDI2_SD)) +
#   geom_point()

table <- df |> 
  select(Participant, BDI2_SD, BDI2_Duration, STAI5_SD, STAI5_Duration) |> 
  mutate_at(vars(ends_with("Duration")), \(x) ifelse(x > 10, 10, x)) |>
  mutate(Participant = fct_reorder(Participant, outliers$Duration)) |> 
  arrange(Participant)

data.frame(Participant = c("Average"), t(sapply(table[2:ncol(table)], mean, na.rm = TRUE))) |>
  rbind(table) |> 
  gt::gt() |> 
  gt::fmt_number() |> 
  # gt::data_color(columns = "Participant", fn=\(x) ifelse(x %in% outliers_half, "orange", "white")) |>
  gt::data_color(columns = ends_with("Duration"), 
                 direction="column", palette = "RdBu") |> 
  gt::data_color(columns = ends_with("SD"), 
                 direction="column", palette = "RdBu", reverse=TRUE) |> 
  gt::data_color(direction="row", rows=1, palette = "grey") |> 
  gt::opt_interactive(use_compact_mode = TRUE)
```

### Attentional Checks

```{r}
df |> 
  select(starts_with("IAS_"), -IAS_Duration, -IAS_Order) |>
  pivot_longer(starts_with("IAS_"), names_to = "item", values_to = "score") |> 
  ggplot(aes(x = score)) +
  geom_density() +
  facet_wrap(~item)
```

## Final Sample

### Participants

The final sample includes `r report::report_participants(df, age="Age", gender="Gender", education="Education")`.

### Age

```{r}
p_age <- df |>  
  mutate(Gender = ifelse(is.na(Gender), "Missing", Gender)) |>
  ggplot(aes(x = x)) +
  geom_histogram(aes(x = Age, fill=Gender), bins=28) +
  geom_vline(xintercept = mean(df$Age), color = "red", linewidth=1.5) +
  # geom_label(data = data.frame(x = mean(df$Age) * 1.15, y = 0.95 * 75), aes(y = y), color = "red", label = paste0("Mean = ", format_value(mean(df$Age)))) +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Age", y = "Number of Participants", color = NULL, subtitle = "Distribution of participants' age") +
  theme_modern(axis.title.space = 10) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1)),
    axis.title.x = element_blank()
  )
p_age
```

<!-- ## Ethnicity -->

<!-- ```{r p_demographics_ethnicity} -->
<!-- plot_waffle <- function(sub, what = "Nationality", title = what, rows = 8, size = 3) { -->
<!--   ggwaffle::waffle_iron(sub, what, rows = rows) |> -->
<!--     ggplot(aes(x, y)) + -->
<!--     geom_point(aes(color = group), shape = "square", size = size) + -->
<!--     coord_equal() + -->
<!--     ggtitle(title) + -->
<!--     labs(fill = "", color = "", subtitle = "N participants per self-declared group") + # (each square represents one participant)" -->
<!--     theme_void() + -->
<!--     theme( -->
<!--       plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0), -->
<!--       plot.subtitle = element_text(size = rel(1.2)), -->
<!--       legend.key.height = unit(1, "mm"), -->
<!--       legend.key.width = unit(1, "mm") -->
<!--     ) -->
<!-- } -->

<!-- p_ethni <- plot_waffle(sub, "Ethnicity", rows = 20, size = 8) + -->
<!--   scale_color_manual(values = c("Hispanic" = "#FF5722", "Caucasian" = "#2196F3", "African" = "#4CAF50", "Asian" = "#FFC107", "Other" = "#795548")) -->
<!-- p_ethni -->
<!-- ``` -->

### Education

```{r}
p_edu <- df |>
  mutate(
    Education = ifelse(Education == "", NA, Education),
    Education = fct_relevel(Education, "Other", "High School", "Bachelor", "Master", "Doctorate")) |> 
  ggplot(aes(x = Education)) +
  geom_bar(aes(fill = Education)) +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Education", y = "Number of Participants", subtitle = "Participants per achieved education level") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1)),
    axis.title.x = element_blank()
  )
p_edu
```

### Table 1

```{r}
table1 <- df |> 
  select(Age, 
         PHQ4_Condition, 
         PHQ4 = PHQ4_Total,
         STAI5 = STAI5_General,
         BDI2 = BDI2_Total) |> 
  report::report_sample(group_by = "PHQ4_Condition")

table1$Variable <- str_remove_all(table1$Variable, fixed("Mean "))
table1$Variable <- str_remove_all(table1$Variable, fixed(" (SD)"))

insight::display(table1)
```

```{r}
add_bf <- function(r) {
  x <- as.data.frame(r)
  f <- as.formula(paste(str_remove(x$Parameter, fixed("df$")), "~ PHQ4_Condition"))
  x$BF <- as.data.frame(BayesFactor::ttestBF(formula=f, data=df))$bf
  x
}

rbind(
  add_bf(report::report(t.test(df$Age ~ df$PHQ4_Condition))),
  add_bf(report::report(t.test(df$PHQ4_Total ~ df$PHQ4_Condition))),
  add_bf(report::report(t.test(df$STAI5_General ~ df$PHQ4_Condition))),
  add_bf(report::report(t.test(df$BDI2_Total ~ df$PHQ4_Condition)))
) |> 
  select(Parameter, t, df_error, p, BF) |> 
  mutate(Parameter = str_remove_all(Parameter, fixed("df$"))) |>
  insight::display()
```

### Distributions

```{r}
df |> 
  select(PHQ4_Condition, 
         STAI5 = STAI5_General,
         BDI2 = BDI2_Total,
         PHQ4_Total) |> 
  standardize() |> 
  estimate_density(at = "PHQ4_Condition") |> 
  ggplot(aes(x=x, y=y)) +
  geom_line(aes(color = Parameter), linewidth=2) +
  scale_color_manual(values = c("STAI5" = "#E91E63", "BDI2" = "#2196F3", "PHQ4_Total" = "#9C27B0")) +
  labs(x = "Standardized Score", y = "Distribution") +
  facet_wrap(~PHQ4_Condition) +
  theme_bw()
```

### Mood Disorders

### Diagnostic

```{r}
p1 <- select(df, Participant, Gender, starts_with("Disorder_")) |> 
  pivot_longer(cols = starts_with("Disorder_"), names_to = "Disorder", values_to = "Value") |> 
  mutate(Disorder = str_remove_all(Disorder, fixed("Disorder_"))) |> 
  summarize(N = sum(Value) / nrow(df), .by=c("Gender", "Disorder")) |> 
  mutate(N_tot = sum(N), .by="Disorder") |>
  mutate(Disorder = fct_reorder(Disorder, desc(N_tot))) |>
  ggplot(aes(x = Disorder, y = N, fill=Gender)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  scale_y_continuous(expand = c(0, 0), labels=scales::percent) +
  labs(title = "Mood Disorders", y = "Number of Participants", subtitle = "Participants per self-declared mood disorder", x="Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p2 <- select(df, Participant, Gender, starts_with("DisorderTreatment_")) |> 
  pivot_longer(cols = starts_with("DisorderTreatment_"), names_to = "Treatment", values_to = "Value") |> 
  mutate(Treatment = str_remove_all(Treatment, fixed("DisorderTreatment_"))) |> 
  summarize(N = sum(Value) / nrow(df), .by=c("Gender", "Treatment")) |> 
  mutate(N_tot = sum(N), .by="Treatment") |>
  mutate(Treatment = fct_reorder(Treatment, desc(N_tot))) |>
  ggplot(aes(x = Treatment, y = N, fill=Gender)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  scale_y_continuous(expand = c(0, 0), labels=scales::percent) +
  labs(title = "Treatment", y = "Number of Participants", subtitle = "Participants per treatment", x="Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p1 / p2
```

```{r}
df$MoodDisorder <- rowSums(select(df, DisorderTreatment_MoodStabilizer, DisorderTreatment_Antidepressant, DisorderTreatment_Anxiolytic))
df$MoodDisorder <- ifelse(df$MoodDisorder > 0, TRUE, FALSE)
df$DisorderHistory <- ifelse(df$DisorderHistory == "Yes", TRUE, FALSE)
```

`r sum(df$MoodDisorder)` participants (`r insight::format_percent(sum(df$MoodDisorder) / nrow(df))`) have a mood disorder, as indexed by the use of a mood stabilizer, antidepressant, or anxiolytic treatment. `r sum(df$DisorderHistory)` participants (`r insight::format_percent(sum(df$DisorderHistory) / nrow(df))`) reported having a history of mood disorder.

## Save

```{r}
#| code-fold: false

write.csv(df, "../data/data.csv", row.names = FALSE)
```